---
title: "Figures for third paper mk2"
author: "Science"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    number_sections: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      dpi = 300,
                      fig.ext = c("png"),
                      fig.width = 10,
                      fig.height = 10,
                      fig.path = "figures/")

options(dplyr.summarise.inform = FALSE)
```



# Brief description

Models were rerun for measles, rubella, meningitis A, Hepatitis B, HPV and yellow fever. Four scenarios were run: no-vaccination, default-update, default-update-catchup, default-nocovid. The impacts were calculated generally as:

default_nocovid = novax - default_nocovid,

default_update = novax - default_update,

default_catchup = default_update - default_update_catchup,

but we also consider default_nocovid-default_update. The time period we consider is 2020 to 2030.

# Where has the 2020-21 disruption left us?

```{r hdw_cross_deaths}
hdw <- headline_diff_full_who()

hdw %>%
  flextable() %>%
  set_caption("Excess deaths per calendar year.")

hdw %>% write.csv("excess_cross_deaths.csv", row.names = FALSE)

```

```{r hdw_other}
hdw <- headline_diff_full_who(outc = "cases")

hdw %>% write.csv("excess_cross_cases.csv", row.names = FALSE)

hdw <- headline_diff_full_who(outc = "dalys")

hdw %>% write.csv("excess_cross_dalys.csv", row.names = FALSE)

hdw <- headline_diff_full_who(d_def = d_def_cohort_2030, d_noc = d_noc_cohort_2030)

hdw %>% write.csv("excess_cohort_deaths.csv", row.names = FALSE)

hdw <- headline_diff_full_who(d_def = d_def_cohort_2030, d_noc = d_noc_cohort_2030, outc = "cases")

hdw %>% write.csv("excess_cohort_cases.csv", row.names = FALSE)

hdw <- headline_diff_full_who(d_def = d_def_cohort_2030, d_noc = d_noc_cohort_2030, outc = "dalys")

hdw %>% write.csv("excess_cohort_dalys.csv", row.names = FALSE)

```

# What can be done to mitigate losses in coverage 

## Headline figures {.tabset}

### Tables

```{r mitigate_headline}
t <- mitigate_table(d_def_cross_2030, grouping = NULL)
t_death <- t %>% filter(outcome == "deaths_catchup_impact") %>% mutate(across(-outcome, round))
out <- paste0(t_death$mean_out, " (95%[", t_death$lo_95, ", ", t_death$hi_95,
              "], 80%[", t_death$lo_80, ", ",t_death$hi_80, 
              "], 50%[", t_death$lo_50, ", ", t_death$hi_50, "]) med =", t_death$median_out)

t2 <- mitigate_table(d_def_cohort_2030, grouping = NULL)
t2_death <- t2 %>% filter(outcome == "deaths_catchup_impact") %>% mutate(across(-outcome, round))
out2 <- paste0(t2_death$mean_out, " (95%[", t2_death$lo_95, ", ", t2_death$hi_95,
              "], 80%[", t2_death$lo_80, ", ",t2_death$hi_80, 
              "], 50%[", t2_death$lo_50, ", ", t2_death$hi_50, "]) med =", t2_death$median_out)
```

Catch-up activities as we have modelled them would avert `r out` deaths between calendar years 2020-2030 and `r out2` between birth cohorts 2020-2030.


```{r mitigate_pc}
noc <- d_noc_cross_2030 %>%
  select(run_id, country, year, deaths_default, modelling_group, disease, who_region) %>%
  rename(deaths_nocovid = deaths_default)

def <- d_def_cross_2030 %>%
  select(run_id, country, year, deaths_default, deaths_catchup, modelling_group, disease, who_region) %>%
  rename(deaths_covid = deaths_default,
         deaths_catchup = deaths_catchup)

comb <- left_join(def, noc, by = join_by(run_id, country, year, modelling_group, disease, who_region))

comb <- comb %>% filter(year %in% 2020:2030)

comb <- comb %>%
  mutate(death_diff = deaths_covid - deaths_nocovid,
         deaths_catch = deaths_covid - deaths_catchup)

comb <- comb %>%
  group_by(disease, run_id, who_region, year, country) %>%
  summarise(death_diff = mean(death_diff, na.rm = TRUE),
            deaths_catch = mean(deaths_catch, na.rm = TRUE))

w <- comb %>%
  group_by(run_id) %>%
  summarise(death_diff = sum(death_diff),
            deaths_catch = sum(deaths_catch)) %>%
  mutate(propn = deaths_catch/death_diff)

w_20_30 <- paste0(round(mean(w$propn*100), 2),  "% (95%[", round(quantile(w$propn*100, 0.025),2), "%, ", round(quantile(w$propn*100, 0.975),2), "%]) med=", round(median(w$propn*100), 2))

num_20_30 <- paste0(round(mean(w$deaths_catch)),  " (95%[", round(quantile(w$deaths_catch, 0.025)), ", ", round(quantile(w$deaths_catch, 0.975)), "]) med=", round(median(w$deaths_catch)))

denom_20_30 <- paste0(round(mean(w$death_diff)),  " (95%[", round(quantile(w$death_diff, 0.025)), ", ", round(quantile(w$death_diff, 0.975)), "]) med=", round(median(w$death_diff)))


comb <- comb %>%
  filter(year %in% 2023:2030)

w <- comb %>%
  group_by(run_id) %>%
  summarise(death_diff = sum(death_diff),
            deaths_catch = sum(deaths_catch)) %>%
  mutate(propn = deaths_catch/death_diff)

w_23_30 <- paste0(round(mean(w$propn*100), 2),  "% (95%[", round(quantile(w$propn*100, 0.025),2), "%, ", round(quantile(w$propn*100, 0.975),2), "%]) med=", round(median(w$propn*100), 2))

num_23_30 <- paste0(round(mean(w$deaths_catch)),  " (95%[", round(quantile(w$deaths_catch, 0.025)), ", ", round(quantile(w$deaths_catch, 0.975)), "]) med=", round(median(w$deaths_catch)))

denom_23_30 <- paste0(round(mean(w$death_diff)),  " (95%[", round(quantile(w$death_diff, 0.025)), ", ", round(quantile(w$death_diff, 0.975)), "]) med=", round(median(w$death_diff)))
```

This represents `r w_20_30` of excess deaths between calendar years 2020 and 2030 [numerator `r num_20_30`, denominator `r denom_20_30`] due to disruptions or `r w_23_30` of excess deaths between calendar years 2023 and 2030 [numerator `r num_23_30`, denominator `r denom_23_30`]. 

```{r mitigate_who_cross}
main <- mitigate_who_table(d_def_cross_2030, tot=t)
main_cases <- mitigate_who_table(d_def_cross_2030, outc = "cases", tot=t)
main_dalys <- mitigate_who_table(d_def_cross_2030, outc = "dalys", tot=t)

main %>%
  flextable(cwidth = 2) %>%
  flextable::set_caption("Mitigated deaths due to catch-up activities- calendar view.")

write.csv(main, "mitigate_cross_deaths.csv", row.names = FALSE)
write.csv(main_cases, "mitigate_cross_cases.csv", row.names = FALSE)
write.csv(main_dalys, "mitigate_cross_dalys.csv", row.names = FALSE)
```

```{r mitigate_who_cohort}
main <- mitigate_who_table(d_def_cohort_2030, tot=t2)
main_cases <- mitigate_who_table(d_def_cohort_2030, outc = "cases", tot=t2)
main_dalys <- mitigate_who_table(d_def_cohort_2030, outc = "dalys", tot=t2)

main %>%
  flextable(cwidth = 2) %>%
  flextable::set_caption("Mitigated deaths due to catch-up activities- cohort view.")

write.csv(main, "mitigate_cohort_deaths.csv", row.names = FALSE)
write.csv(main_cases, "mitigate_cohort_cases.csv", row.names = FALSE)
write.csv(main_dalys, "mitigate_cohort_dalys.csv", row.names = FALSE)
```

# Full uncertainty figures {.tabset}

## cross

```{r burden_unc}
noc <- d_noc_cross_2030 %>%
  select(run_id, country, year, deaths_default,cases_default, dalys_default, modelling_group, disease, who_region) %>%
  rename(deaths_nocovid = deaths_default,
         cases_nocovid = cases_default, 
         dalys_nocovid = dalys_default)

def <- d_def_cross_2030 %>%
  select(run_id, country, year, deaths_default, cases_default, dalys_default, modelling_group, disease, who_region) %>%
  rename(deaths_covid = deaths_default,
         cases_covid = cases_default,
         dalys_covid = dalys_default)

comb <- left_join(def, noc, by = join_by(run_id, country, year, modelling_group, disease, who_region))

comb <- comb %>%
  mutate(death_diff = deaths_covid-deaths_nocovid,
         case_diff  = cases_covid - cases_nocovid,
         daly_diff = dalys_covid - dalys_nocovid)

comb <- comb %>%
  group_by(run_id, country, year, who_region, disease) %>%
  summarise(death_diff = mean(death_diff),
            deaths_covid = mean(deaths_covid),
            deaths_nocovid = mean(deaths_nocovid),
            case_diff = mean(case_diff),
            cases_covid = mean(cases_covid),
            cases_nocovid = mean(cases_nocovid),
            daly_diff = mean(daly_diff),
            dalys_covid = mean(dalys_covid),
            dalys_nocovid = mean(dalys_nocovid),)


comb3 <- comb %>%
  ungroup() %>%
  select(run_id, country, year, disease, deaths_covid, deaths_nocovid) %>%
  pivot_longer(names_to = "scenario", values_to = "deaths", -c(run_id, country, year, disease)) %>%
  group_by(run_id, year, disease, scenario) %>%
  summarise(deaths = sum(deaths))


```



```{r excess_deaths_full_uncert_cross ,fig.cap="Highlighted if median is greater than 200 at any point, only figures where the abs(median) excess is >50 are included."}
p <- fun_excess_ribbon_95(comb, "deaths", thresh=200, thresh2=50)

ggsave(plot=p,  filename="excess_deaths_full_uncert_cross-1.png", path="figures", bg = "white", width = 12, height = 10)
ggsave(plot=p,  filename="excess_deaths_full_uncert_cross-1.pdf", path="figures", bg = "white", width = 12, height = 10)
```

```{r excess_cases_full_uncert_cross, fig.cap="Highlighted if median is greater than 200 at any point, only figures where the abs(median) excess is >50 are included."}
fun_excess_ribbon(comb, "cases", thresh=200, thresh2=50)
```

```{r excess_dalys_full_uncert_cross, fig.cap="Highlighted if median is greater than 200 at any point, only figures where the abs(median) excess is >50 are included."}
fun_excess_ribbon(comb, "dalys", thresh=200, thresh2=50)
```


```{r cross_solar_system_who, fig.width = 12}

fun_solar_system()
```


```{r cross_solar_system_who_2023, fig.width = 12}

fun_solar_system(yar = 2023:2030)
```

```{r cross_solar_system_who_2023_determ, fig.width = 12}

p <- fun_solar_system(yar = 2023:2030, determ=TRUE)
ggsave(plot=p,  filename="cross_solar_system_who_2023_determ-1.png", path="figures", bg = "white", width = 12, height = 10)
ggsave(plot=p,  filename="cross_solar_system_who_2023_determ-1.pdf", path="figures", bg = "white", width = 12, height = 10)
```

```{r cross_solar_system_who_2023_cases, fig.width = 12}

fun_solar_system(yar = 2023:2030, outc = "cases")
```

```{r cross_solar_system_who_2023_cases_determ, fig.width = 12}

fun_solar_system(yar = 2023:2030, outc = "cases", determ=TRUE)
```

```{r cross_solar_system_who_2023_dalys, fig.width = 12}

fun_solar_system(yar = 2023:2030, outc="dalys")
```

```{r cross_solar_system_who_2023_dalys_determ, fig.width = 12}

fun_solar_system(yar = 2023:2030, outc="dalys", determ=TRUE)
```

```{r plot_mitigate_timing_cross_deaths, fig.width=12, fig.height=12}
p <- plot_mitigate_timing_95(outc = "deaths", thresh2=50)

ggsave(plot=p,  filename="plot_mitigate_timing_cross_deaths-1.png", path="figures", bg = "white", width = 12, height = 10)
ggsave(plot=p,  filename="plot_mitigate_timing_cross_deaths-1.pdf", path="figures", bg = "white", width = 12, height = 10)
```

```{r plot_mitigate_timing_cross_cases, fig.width=12, fig.height=12}
plot_mitigate_timing(outc = "cases", thresh2=50)
```

```{r plot_mitigate_timing_cross_dalys, fig.width=12, fig.height=12}
plot_mitigate_timing(outc = "dalys", thresh2=50)
```

# Interim update
In this section we perform a simple interim update on the `covidimpactiu` touchstone. 

```{r agg data}

comb_dat <- bind_rows(lapply(c(1), 
                             function(x){
                               fun_join_fvps_df_iu(fvps, df, RI_scaling = x, ctry_subset = ctry_subset_parm, dis_subset = NULL) %>%
                                 mutate(scaling = x)
                             }))

cc <- make_country_classifiers(con)

# aggregate
comb_dat_agg_ave <- comb_dat %>%
  group_by(country, disease, vaccine, activity_type,  year, scenario_description, scenario_type, scaling, run_id) %>%
  summarise(deaths = sum(deaths, na.rm = TRUE), cases = sum(cases, na.rm = TRUE), dalys = sum(dalys, na.rm = TRUE))


```

## Overall differences {.tabset}

We present the percentage difference to the `default-update` baseline over the entire time period 2020:2030 for each scaling of the potential routine-intensified impact ratio.

```{r diff}
comb_dat_agg_ave <- comb_dat_agg_ave %>%
  left_join(comb_dat_agg_ave %>%
              filter(scenario_type %in% "default_update") %>% 
              ungroup() %>%
              select(-c(scenario_type, scenario_description)) %>% 
              rename(deaths_baseline = deaths, cases_baseline = cases, dalys_baseline = dalys),
            by = c("country", "disease", "vaccine", "activity_type", "year", "scaling", "run_id"))

comb_dat_agg_ave <- comb_dat_agg_ave %>% left_join(cc, by = "country")
```


## What is the breakdown of excess deaths by year and activity type? {.tabset}


```{r }
t <- comb_dat_agg_ave %>%
  filter(scenario_type %in% c("default_nocovid", "default_update")) %>%
  filter(scaling == 1) %>%
  filter(year %in% 2020:2030) %>%
  group_by(scenario_type, run_id) %>%
  summarise(all_deaths_averted = sum(deaths, na.rm = TRUE)) %>%
  pivot_wider(names_from = scenario_type, values_from = all_deaths_averted)


noc <- tidy_qi(t$default_nocovid)
noc10 <- tidy_qi(t$default_nocovid/10)
cov <- tidy_qi(t$default_update)
cov10 <- tidy_qi(t$default_update/10)

diff <- tidy_qi(t$default_nocovid-t$default_update)
diffpc <- tidy_qi((t$default_nocovid-t$default_update)/t$default_update*100, roundamt = 2)
```

We estimate that in the absence of COVID-related disruptions, with conservative targets by 2030, there will be `r noc` deaths averted over the decade, so `r noc10` on average a year. In contrast, with COVID-related disruptions and recovery, we estimate that `r cov` will be averted over the decade, so `r cov10` on average a year. This is a difference of `r diffpc`%. We may also consider this by country, below.

```{r }
t <- comb_dat_agg_ave %>%
  filter(scenario_type %in% c("default_nocovid", "default_update")) %>%
  filter(scaling == 1) %>%
  filter(year %in% 2020:2030) %>%
  group_by(scenario_type, run_id, country) %>%
  summarise(all_deaths_averted = sum(deaths, na.rm = TRUE)) %>%
  pivot_wider(names_from = scenario_type, values_from = all_deaths_averted)

t2 <- t

t2 <- t2 %>%
  group_by(country) %>%
  summarise(noc = tidy_qi(default_nocovid),
            cov = tidy_qi(default_update),
            diff = tidy_qi(default_nocovid - default_update),
            diffpc = tidy_qi((default_nocovid-default_update)/default_update*100))

t2 %>% flextable(cwidth=2)

t2 %>% write.csv("pc_drop_per_country.csv", row.names = FALSE)
```



```{r }
t <- comb_dat_agg_ave %>%
  filter(scenario_type %in% c("default_nocovid", "default_update")) %>%
  filter(scaling == 1) %>%
  filter(year %in% 2020:2030) %>%
  group_by(scenario_type, activity_type, run_id) %>%
  summarise(all_deaths_averted = sum(deaths, na.rm = TRUE)) %>%
  pivot_wider(names_from = scenario_type, values_from = all_deaths_averted) %>%
  mutate(diff = default_nocovid - default_update) %>%
  group_by(run_id) %>%
  mutate(propn = diff/sum(diff))

campaign_prop <- tidy_qi(t$propn[t$activity_type=="campaign"]*100, 2)
routine_prop <- tidy_qi(t$propn[t$activity_type=="routine"]*100, 2)
```

Of the deaths averted over the decade, and the excess deaths due to coverage disruptions, `r campaign_prop`% is due to differences in campaigns and `r routine_prop`% is due to routine differences. The breakdown is shown below.

```{r}
t <- comb_dat_agg_ave %>%
  filter(scenario_type %in% c("default_nocovid", "default_update")) %>%
  filter(scaling == 1) %>%
  filter(year %in% 2020:2030) %>%
  group_by(scenario_type, activity_type, disease, run_id) %>%

  summarise(all_deaths_averted = sum(deaths, na.rm = TRUE)) %>%
  pivot_wider(names_from = scenario_type, values_from = all_deaths_averted) %>%
  mutate(diff = default_nocovid - default_update) %>%
  group_by(disease, run_id) %>%
  mutate(propn = diff/sum(diff)) %>% arrange(disease) 



t %>% 
  group_by(disease, activity_type) %>% 
  summarise(propn = tidy_qi(propn, roundamt = 2)) %>%
  flextable(cwidth = 2)

write.csv(t%>% 
  group_by(disease, activity_type) %>% 
  summarise(propn = tidy_qi(propn, roundamt = 4)), "difference_activity_type.csv", row.names = FALSE)
```

```{r}
t <- comb_dat_agg_ave %>%
  filter(scenario_type %in% c("default_nocovid", "default_update")) %>%
  filter(scaling == 1) %>%
  filter(year %in% 2020:2030) %>%
  group_by(scenario_type, year, disease, run_id) %>%

  summarise(all_deaths_averted = sum(deaths, na.rm = TRUE)) %>%
  pivot_wider(names_from = scenario_type, values_from = all_deaths_averted) %>%
  mutate(diff = default_nocovid - default_update) %>%
  group_by(disease) %>%
  mutate(propn = diff/sum(diff)) %>% arrange(disease) %>%
  mutate(year_group = ifelse(year %in% 2020:2021, "2020-1", ifelse(year %in% 2022:2025, "2022-5", "2026-30"))) %>%
  group_by(year_group, disease, run_id) %>%
  summarise(diff_group = round(sum(diff))) %>%
  group_by(disease, run_id) %>%
  mutate(propn_group = round(diff_group/sum(diff_group)*100, 2))

t %>% mutate(year_group = as.character(year_group)) %>% 
  group_by(disease, year_group) %>% 
  summarise(propn_group = tidy_qi(propn_group, roundamt = 2)) %>%
  flextable(cwidth = 2)

write.csv(t%>% mutate(year_group = as.character(year_group)) %>% 
  group_by(disease, year_group) %>% 
  summarise(propn_group = tidy_qi(propn_group, roundamt = 4)), "difference_year.csv", row.names = FALSE)
```

# All disease stacked bar

```{r disease_stacked_bar, fig.height=8, fig.width=12, fig.cap="Interim update projected excess deaths (hundreds) due to coverage disruption between 2020 and 2030"}
yar <- 2020:2030
df_iu <- df_iu %>% left_join(ctry %>% select(country, who_region))

t <- df_iu %>%
  rename(short_name = scenario_type, deaths_averted = deaths) %>%
  filter(year %in% yar) %>%
  group_by(disease, year, short_name, country, who_region, vaccine) %>%
  mutate(who_region = ifelse(who_region=="PAHO", who_region, gsub("O", "", who_region))) %>%
  summarise(deaths_averted = mean(deaths_averted, na.rm = TRUE)) %>%
  group_by(disease, short_name, who_region) %>%
  summarise(deaths_averted = sum(deaths_averted, na.rm = TRUE)) %>%
  select(disease, short_name, deaths_averted, who_region) %>%
  pivot_wider(names_from = short_name, values_from = deaths_averted) %>%
  mutate(excess_deaths = default_nocovid - default_update) %>%
  mutate(disease = as.factor(disease))%>%
  filter((excess_deaths)>1)%>%
  left_join(ctry %>% group_by(who_region) %>% summarise(n_ctry = length(unique(country))), by = "who_region") 

t <- t %>%
  mutate(group = disease, value = excess_deaths/100, who_region = paste0(who_region, " (", n_ctry," VIMC countries)")) 

p <- t %>% 
  ggplot()+
  aes(fill = group, y = value, x = who_region)+
  geom_col()+
  scale_fill_manual(values=disease_palette(unique(t$disease))$pal)+
  coord_flip()+
  theme_minimal()+
  #facet_wrap(who_region~., ncol = 1)+
  theme(legend.position = "bottom")+
  labs(fill = "Disease/ Vaccine", x = "WHO region", y = "Excess deaths (100s)")

ggsave(plot=p,  filename="disease_stacked_bar-1.pdf", path="figures", bg = "white", width = 12, height = 8)

```